02_format_and_clean

### 2022-07-17 replication_data/rmd/02_format_and_clean.rmd
v1.1

`%>%` <- magrittr::`%>%`
```

# Load and transform wide to long

read in icb_coder_crisis_sentence_event_aligned.Rds
```{r}
icb_wide_aligned_raw <- readRDS(file=paste0(here::here(),"/replication_data/temp/ICBe_V1_long_unclean.Rds")) %>% 
                    dplyr::mutate(crisno3 = stringr::str_pad(crisno, 3, pad = "0"), sentence_number_int_aligned3 =
                                    stringr::str_pad(sentence_number_int_aligned, 3, pad = "0")) %>%
                    dplyr::mutate(email_crisis_sent_event =
                                    paste0(email_id,"_",crisno3,"_",sentence_number_int_aligned3,"_",event_number_int)) %>%
                    dplyr::select(-crisno3, -sentence_number_int_aligned3) %>%
                    dplyr::arrange(crisno, sentence_number_int, event_number_int) %>%

                    ####################Where we kill unhelpful events
                    #There are 2 legitimate kinds of rows
                    dplyr::filter(
                      event_number_int==0 & sentence_events %in% c('Background', 'Covered by Previous Sentence','Error','No Action') |
                      event_number_int>0 & event_type %in% c('Action', 'Speech','Thought' )
                    ) #%>%

dim(icb_wide_aligned_raw) #87710    85
```

Remove variables used earlier for compiling the individual codings
```{r}
icb_wide_aligned <- icb_wide_aligned_raw %>% dplyr::ungroup() %>% 
  dplyr::select(
               # -lastviewed, 
               # -input_crisis, 
                -sentence_number_int, -raw_crisis, -raw_crisno, -raw_sentence, -savefile_crc32, -raw_crisno_sentno, -raw_crisno_version_sentno, -crisisno_crisis_version) %>%
  dplyr::mutate(icb_survey_version = ifelse(is.na(icb_survey_version), 1.1, icb_survey_version))
dim(icb_wide_aligned) #52929    78

@@

Initial coding edits (spelling, formatting, etc)
```{r}
icb_wide_aligned <- icb_wide_aligned %>% 
                      dplyr::rename(act_deescalate = act_descalate) %>% 
                      dplyr::rename(interact_deescalate = interact_descalate) %>% 
                      dplyr::rename(condition_act_deescalate = condition_act_descalate) %>% 
                      dplyr::rename(condition_interact_deescalate = condition_interact_descalate) %>% 
                      dplyr::mutate(sayintkind_react = ifelse(sayintkind_react=="This Sentence", raw_sentence_number_int,
                                                                         stringr::str_replace(sayintkind_react, "Sentence ","")))
dim(icb_wide_aligned) #52929    78

@@

Where a coder looked at the same crisis twice, drop the earlier version and the earlier date
```{r}
icb_wide_aligned <- icb_wide_aligned %>% 
                      dplyr::group_by(email_crisis_sent_event) %>% 
                      dplyr::slice_max(icb_survey_version) %>%
                      dplyr::ungroup()

icb_wide_aligned <- icb_wide_aligned %>% 
                    dplyr::group_by(email_crisis_sent_event) %>% 
                    dplyr::slice_max(lastviewed) %>%
                    dplyr::ungroup()
dim(icb_wide_aligned) #87710    77
```

Drop a small subset where it appears that a coder looked at the same crisis twice but one of the times was logged incorrectly
```{r}
icb_wide_aligned <- icb_wide_aligned %>% dplyr::left_join(
                    icb_wide_aligned %>% dplyr::select(email_crisis_sent_event, event_number_int) %>% dplyr::filter(
                    (icb_wide_aligned %>% dplyr::select(email_crisis_sent_event, event_number_int) %>% 
                       duplicated()) == TRUE) %>% 
                      dplyr::mutate(duplicate = 1)) %>% 
                    dplyr::mutate(mismatch = crisis != input_crisis) %>% 
                    dplyr::filter(!(mismatch %in% TRUE & duplicate %in% 1)) %>%
                    dplyr::select(-mismatch, -duplicate)

dim(icb_wide_aligned) #51098    78

@@

Pivot to long form 
```{r}

icb_long <- icb_wide_aligned %>% dplyr::ungroup() %>% 
             dplyr::relocate(sentence_events, .after = event_number_int) %>%
             tidyr::pivot_longer(
               cols = c(sentence_events:interact_units), #  changed this to be inclusive of event type and sentence_events 
               names_to = "variable",
               #names_prefix = "wk",
               values_to = "value",
               values_drop_na = TRUE
             ) %>% dplyr::ungroup() %>% 
  dplyr::mutate(value=value %>% tolower %>% trimws())
dim(icb_long) #491757     15

Now split values with a list (separated by a semicolon)

icb_long_split <- icb_long %>% 
                  dplyr::mutate(value = strsplit(as.character(value), ";")) %>% 
                  tidyr::unnest(value) %>%
                  dplyr::mutate(value=value %>% tolower %>% trimws())
dim(icb_long_split) #536397     15

# Define subsets of variables

vars <- c(
"sentence_events", "event_type", "raterconfidence", "raterconfidence_reason",
"raterconfidence_reason_survey", "date", "think_sentence_events", "think_actor_a", "think_actor_a_other", 
"thinkkind", "say_actor_a", "say_actor_a_other", "say_actor_b", "say_actor_b_other",
"sayintkind", "sayintkind_react", "consequence", "condition", "do_actor_a",
 "do_actor_a_other", "do_actor_b", "do_actor_b_other", "condition_do_actor_a",
"condition_do_actor_a_other", "condition_do_actor_b", "condition_do_actor_b_other",
 "condition_act_cooperative", "condition_act_deescalate", "condition_act_escalate",
"condition_act_uncooperative", "condition_do_interact_kind", "condition_do_kind",
"condition_interact_decreasecoop",  "condition_interact_deescalate",
"condition_interact_domains", "condition_interact_escalate", "condition_interact_fatalities",
"condition_interact_forces", "condition_interact_geoscope", "condition_interact_increasecoop",
"condition_interact_location", "condition_interact_location_other", "condition_interact_territory",
"condition_interact_units", "do_duration", "do_interact_kind", "do_kind", "do_timing",
"do_timing_reorder", "act_cooperative", "act_deescalate", "act_escalate",
"act_uncooperative", "interact_decreasecoop", "interact_deescalate", "interact_domains",
"interact_escalate", "interact_fatalities", "interact_forces", "interact_geoscope",
"interact_increasecoop",  "interact_location", "interact_location_other", "interact_territory",
"interact_units" )

vars_actor <- c("think_actor_a",
             "think_actor_a_other",
             "say_actor_a",
             "say_actor_a_other",
             "say_actor_b",
             "say_actor_b_other",
             "do_actor_a",
             "do_actor_a_other",
             "do_actor_b",
             "do_actor_b_other", 
             "condition_do_actor_a",
             "condition_do_actor_a_other",
             "condition_do_actor_b",
             "condition_do_actor_b_other")

vars_act <- c("act_cooperative",
              "act_uncooperative",
              "act_deescalate",
              "act_escalate",
              "interact_decreasecoop",
              "interact_increasecoop",
              "interact_deescalate",
              "interact_escalate",
              "condition_act_cooperative",
              "condition_act_uncooperative",
              "condition_act_deescalate",
              "condition_act_escalate",
              "condition_interact_decreasecoop",
              "condition_interact_increasecoop",
              "condition_interact_deescalate",
              "condition_interact_escalate")

# saveRDS(vars_actor, paste0(here::here(), "/replication_data/temp/","vars_actor.Rds"))
# saveRDS(vars_act, paste0(here::here(), "/replication_data/temp/","vars_act.Rds"))

vars_say <- c("sayintkind", "sayintkind_react","consequence","condition")
vars_think <- c("think_sentence_events", "thinkkind", "thinkkind_clean")

# Duplicate variables, same options for each in these types
vars_act_kind <- c("do_kind", "condition_do_kind", "do_interact_kind", "condition_do_interact_kind")
vars_fatalities <- c("interact_fatalities", "condition_interact_fatalities")
vars_forces <- c("interact_forces", "condition_interact_forces")
vars_geoscope <- c("interact_geoscope", "condition_interact_geoscope")
vars_location <- c("interact_location", "condition_interact_location")
vars_location_other <- c("interact_location_other", "condition_interact_location_other")
vars_territory <- c("interact_territory", "condition_interact_territory")

# NEED TO ADD TO CODEBOOK
vars_domains <- c("interact_domains", "condition_interact_domains")
vars_units <- c("interact_units", "condition_interact_units")

```{r}
# are all variables in vars?
setdiff(unique(icb_long_split$variable), vars)

# do the subsets cover everything in vars?
setdiff(vars, c(vars_actor, vars_act, vars_act_kind, vars_fatalities, vars_forces, vars_geoscope, vars_location, vars_location_other, vars_territory, vars_domains, vars_units, vars_say, vars_think))
```

@@

```{r message=FALSE}
target_file <- paste0(here::here(),"/replication_data/in/icb_manual_recoding_master_sheet.xlsx")
# in not working, run googledrive::drive_auth() and select new token

dictionary_actions  <- readxl::read_excel(target_file, sheet="actions") %>% dplyr::distinct()
dictionary_actors    <- readxl::read_excel(target_file, sheet="actors") %>% dplyr::select(crisno,value, value_normalized=value_normalized_wikidata_id) %>% dplyr::distinct()

#Locations and dates are both really hard to do
dictionary_locations <- readxl::read_excel(target_file, sheet="locations", na = c("", "NA")) 
dictionary_dates <- readxl::read_excel(target_file, sheet="dates", col_types="text") #This is the tricky one and why he had to add values

```{r}
# Create dateset to reconnect location dictionary to individual rows through 'sentence_number_int_aligned'
location_merge_directory <- icb_wide_aligned %>% 
    dplyr::mutate(crisno3 = stringr::str_pad(crisno, 3, pad = "0"), 
                  sentence_number_int_aligned3 = stringr::str_pad(sentence_number_int_aligned, 3, pad = "0")) %>% 
  dplyr::mutate(email_crisis_sent_event = paste(email_id,crisno3,sentence_number_int_aligned3,event_number_int, sep="_")) %>%
  dplyr::select(-crisno3, -sentence_number_int_aligned3) %>%
  dplyr::mutate(crisis_sent = paste0(crisno,"_",sentence_number_int_aligned)) %>%
  dplyr::select(crisno, email_crisis_sent_event, do_actor_a, do_actor_b, interact_location, interact_location_other) %>% 
  dplyr::mutate(do_actor_a=strsplit(do_actor_a, ";")) %>% tidyr::unnest(do_actor_a) %>%
  dplyr::mutate(do_actor_b=strsplit(do_actor_b, ";")) %>% tidyr::unnest(do_actor_b) %>%
  dplyr::mutate(interact_location_other=strsplit(interact_location_other, ";")) %>% tidyr::unnest(interact_location_other) %>%
  dplyr::mutate(do_actor_a=tolower(do_actor_a), do_actor_b=tolower(do_actor_b),
                interact_location_other=trimws(tolower(interact_location_other))) %>%
  dplyr::left_join(dictionary_locations %>%
                     dplyr::mutate() %>% 
                     dplyr::mutate(interact_location_other = tolower(interact_location_other)) %>% 
                     dplyr::select(do_actor_a, do_actor_b, interact_location, interact_location_other, norm_name)) %>%
  dplyr::mutate(crisno = as.numeric(crisno)) %>% 
  unique() %>% 
  dplyr::filter(!is.na(interact_location_other))

Apply the cleaning dictionaries to the relevant subsets

```{r}
  #ACTOR
icb_long_actor_clean <- icb_long_split %>% 
  dplyr::filter(variable %in% vars_actor) %>%
    dplyr::left_join(tibble::tibble(variable=vars_actor) %>% 
                       dplyr::mutate(temp="temp") %>%
                       dplyr::left_join(dictionary_actors %>% dplyr::mutate(temp="temp")) %>%
                       dplyr::select(-temp) %>% 
                       dplyr::select(crisno, variable, value, value_normalized) %>%
                     dplyr::mutate(crisno = as.character(crisno)) %>%
                     # dplyr::mutate(value_normalized = ifelse(is.na(value_normalized), "no_cleaning", value_normalized)) %>% 
                       unique(),
                     by=c('crisno','variable','value')) %>%
    dplyr::mutate(variable_normalized = paste0(variable, "_clean")) %>% 
  dplyr::mutate(value_normalized = strsplit(as.character(value_normalized), ";")) %>% 
  tidyr::unnest(value_normalized) %>%
  dplyr::mutate(value_normalized=value_normalized %>% trimws())

# check for unmatched
icb_long_actor_clean %>% dplyr::filter(is.na(value_normalized)) %>% dplyr::select(crisno, variable, value)
```

```{r}
  #ACTION
icb_long_action_clean <- icb_long_split %>% 
                        dplyr::filter(variable %in% vars_act) %>% 
                        dplyr::left_join(dplyr::bind_rows(dictionary_actions,
                                         dictionary_actions %>% 
                                           dplyr::mutate(variable = paste0("condition_", variable)) %>% 
                                           dplyr::mutate(variable_normalized = ifelse(is.na(variable_normalized), 
                                                                                            variable_normalized, 
                                                                                            paste0("condition_", variable_normalized)))) %>% 
                                           dplyr::select(variable, value, variable_normalized, value_normalized) %>%
                                           dplyr::mutate(variable_normalized = ifelse(!is.na(value_normalized) & is.na(variable_normalized), 
                                                                                   variable, variable_normalized)),
                                         by=c('variable','value'))

# check for unmatched
icb_long_action_clean %>% dplyr::filter(is.na(value_normalized)) %>% dplyr::select(crisno, variable, value, variable_normalized, value_normalized) 
```

```{r}
 #LOCATION
icb_long_location_clean <- icb_long_split %>% 
  dplyr::filter(variable == "interact_location_other") %>% 
  dplyr::full_join(location_merge_directory %>%
                     dplyr::mutate(crisno = as.character(crisno)) %>% 
                     dplyr::rename(value = interact_location_other)) %>% 
  dplyr::rename(value_normalized = norm_name) %>%
  dplyr::mutate(variable_normalized = "interact_location_other_clean") %>%   
  dplyr::select(-do_actor_a, -do_actor_b, -interact_location) %>% 
  dplyr::mutate(value_normalized = ifelse(is.na(value_normalized), "no_cleaning", value_normalized))
```

```{r}
  #DATES
icb_long_dates_clean <- icb_long_split %>% 
  dplyr::filter(variable=="date") %>% 
  dplyr::left_join(dictionary_dates %>% 
                   dplyr::select(-coder, -notes) %>%
                   dplyr::rename(value=date_orig)) %>%
  tidyr::pivot_longer(cols = point_in_time:latest_date2,
               names_to = "variable_normalized",
               values_to = "value_normalized",
               values_drop_na = TRUE) %>%
  dplyr::mutate(variable_normalized = ifelse(variable_normalized %in% c("point_in_time2", "point_in_time3", "point_in_time4", "point_in_time5"), "point_in_time", variable_normalized))

```{r}
# combine cleaned variables

icb_long_clean <- icb_long_split %>% 
  dplyr::left_join(icb_long_action_clean, by=names(icb_long_split)) %>% 
  dplyr::left_join(icb_long_actor_clean, by=names(icb_long_split), suffix = c("", ".actor")) %>% 
      dplyr::mutate(variable_normalized = (ifelse(is.na(variable_normalized), variable_normalized.actor, variable_normalized))) %>%
      dplyr::mutate(value_normalized = (ifelse(is.na(value_normalized), value_normalized.actor, value_normalized))) %>%
      dplyr::select(-variable_normalized.actor, -value_normalized.actor) %>% 
  dplyr::left_join(icb_long_location_clean, by=names(icb_long_split), suffix = c("", ".location")) %>% 
      dplyr::mutate(variable_normalized = (ifelse(is.na(variable_normalized), variable_normalized.location, variable_normalized))) %>%
      dplyr::mutate(value_normalized = (ifelse(is.na(value_normalized), value_normalized.location, value_normalized))) %>%
      dplyr::select(-variable_normalized.location, -value_normalized.location) %>% 
  dplyr::full_join(icb_long_dates_clean, by=names(icb_long_split), suffix = c("", ".date")) %>% 
      dplyr::mutate(variable_normalized = (ifelse(is.na(variable_normalized), variable_normalized.date, variable_normalized))) %>%
      dplyr::mutate(value_normalized = (ifelse(is.na(value_normalized), value_normalized.date, value_normalized))) %>%
      dplyr::select(-variable_normalized.date, -value_normalized.date)

```{r}
# As the dictionaries translate some groups into multiple values, need to split out rows with multiple values

icb_long_clean <- icb_long_clean %>% 
  dplyr::mutate(value_normalized = strsplit(as.character(value_normalized), ";")) %>% 
  tidyr::unnest(value_normalized) %>%
  dplyr::mutate(value_normalized=value_normalized %>% trimws()) %>% 
  dplyr::mutate(crisno=as.numeric(crisno)) %>% 
  dplyr::mutate(value=tolower(value))

dim(icb_long_clean) #507699
```

```{r}
# Fix several typos:

print(paste("change 'premption' to 'preemption'", icb_long_clean %>% dplyr::filter(stringr::str_detect(value, "premption")==TRUE) %>% nrow(), "times."))
icb_long_clean <- icb_long_clean %>% dplyr::mutate(value = stringr::str_replace(value, "premption", "preemption"))

print(paste("change 'disaprove' to 'disapprove'", icb_long_clean %>% dplyr::filter(stringr::str_detect(value, "disaprove")==TRUE) %>% nrow(), "times."))
icb_long_clean <- icb_long_clean %>% dplyr::mutate(value = stringr::str_replace(value, "disaprove", "disapprove"))

print(paste("change 'uncoperative' to 'uncooperative'", icb_long_clean %>% dplyr::filter(stringr::str_detect(value, "uncoperative")==TRUE) %>% nrow(), "times."))
icb_long_clean <- icb_long_clean %>% dplyr::mutate(value = stringr::str_replace(value, "uncoperative", "uncooperative"))

print(paste("change 'capitol' to 'capital'", icb_long_clean %>% dplyr::filter(stringr::str_detect(value, "capitol")==TRUE) %>% nrow(), "times."))
icb_long_clean <- icb_long_clean %>% dplyr::mutate(value = stringr::str_replace(value, "capitol", "capital"))

print(paste("change 'survailance' to 'surveillance'", icb_long_clean %>% dplyr::filter(stringr::str_detect(value, "survailance")==TRUE) %>% nrow(), "times."))
icb_long_clean <- icb_long_clean %>% dplyr::mutate(value = stringr::str_replace(value, "survailance", "surveillance"))

>>>>

library(tidyverse)

`%>%` <- magrittr::`%>%`

@@

This assigns sentence numbers by the text spans instead of the original gui

icb_long_spans <- readRDS(file=paste0(here::here(),"/replication_data/in/icb_long_spans.Rds"))

coders_per_crisis <- icb_long_spans %>% dplyr::select(crisno,  email_id, expert) %>% distinct() #%>% dplyr::count(crisno)
coders_per_crisis_tally <- coders_per_crisis %>% mutate(novice=expert==0) %>% group_by(crisno) %>% summarise( expert=sum(expert==1), novice=sum(novice==1) )
crisis_sentences_per_coder <- icb_long_spans %>% filter(varname %>% str_detect("sentence_events")) %>% dplyr::count(email_id, crisno)

dim(icb_long_spans) #767100     16
icb_long_spans_clean <- icb_long_spans %>%
      filter(!is.na(sentence_span_start) & sentence_span_score>0.35 & sentence_clean!='PRE-CRISIS') %>%
      group_by(crisno) %>%
        mutate(sentence_number_int_aligned=sentence_span_start %>% as.factor() %>% as.integer() ) %>%
      ungroup() %>%
      filter(!varname %in% c('sentence_number_int', 'raw_crisis', 'raw_crisno', 'raw_sentence', 'savefile_crc32', 'raw_crisno_sentno', 'raw_crisno_version_sentno', 'crisisno_crisis_version')) %>%
      mutate(varname=varname %>% recode(act_descalate = 'act_deescalate', interact_descalate = 'interact_deescalate', condition_act_descalate = 'condition_act_deescalate', 
                                        condition_interact_descalate = 'condition_interact_deescalate' ) ) %>%
  filter(!varname %in% c("sentence","text")) %>% 
  dplyr::mutate(value = strsplit(as.character(value), ";")) %>% 
  tidyr::unnest(value) %>%
  dplyr::mutate(value=value %>% tolower %>% trimws()) %>% 
  filter(!duplicated(paste(email_id, crisno,  sentence_number_int_aligned, event_number_int, varname,value)))  %>%
  mutate(sentencenumber_original= value %>% as.numeric() ) %>% #do this out of the loop or the warnings slow everything down
  group_by(email_id, crisno, input_crisis,sentence_number_int_aligned ) %>% 
    mutate(sentencenumber_original= ifelse(varname %in%  'sentencenumber', value, NA) ) %>%
    mutate(sentencenumber_original= sentencenumber_original %>% max(na.rm=T)) %>%
  ungroup()

dim(icb_long_spans_clean) #715451

varname_count <- icb_long_spans_clean %>% count(varname)
value_count <- icb_long_spans_clean %>% count(value)

#dplyr::mutate(sayintkind_react = ifelse(sayintkind_react=="This Sentence", raw_sentence_number_int,
#                                                   stringr::str_replace(sayintkind_react, "Sentence ","")))

#we basically don't believe 'sentence  1' that just means earlier

#Map the original sentence reorder to the new span numbers
icb_long_spans_clean %>% filter(varname %in% c('do_timing','do_timing','think_sentence_events','sayintkind_react','do_timing_reorder')) %>% count(varname, value) %>% View()

alter_allignment <- icb_long_spans_clean %>% 
                    dplyr::select(crisno, email_id, input_crisis, sentence_number_int_aligned,sentencenumber_original) %>% distinct()

reorder_normalized <- icb_long_spans_clean %>% 
                      filter(varname %in% c('do_timing','do_timing','think_sentence_events','sayintkind_react','do_timing_reorder')) %>%
                      dplyr::select(crisno, email_id, input_crisis, varname, value,sentence_number_int_aligned,sentencenumber_original) %>%
                      mutate(value_normalized = value %>% str_replace("sentence  |sentence ","") %>% trimws() %>% as.numeric() ) %>%
                      left_join(
                        alter_allignment %>% dplyr::select(crisno , email_id ,input_crisis  ,value_normalized_new=sentence_number_int_aligned ,value_normalized=sentencenumber_original) %>%
                          mutate(value_normalized=value_normalized %>% as.numeric(), value_normalized_new= value_normalized_new %>% as.numeric() )
                        ) %>%
                      mutate(value_normalized = ifelse(!is.na(value_normalized_new) , value_normalized_new, value_normalized) ) %>%
                      dplyr::select(-value_normalized_new) %>%
                      mutate(value_normalized = ifelse(value %in% c("act in this sentence","speech in this sentence","this sentence") , sentence_number_int_aligned, value_normalized) ) %>%
                      mutate(value_normalized=value_normalized %>% as.character()) %>%
                      #reorders were broken and set to default sentence 1 so drop them without loss of info
                      mutate(value_normalized = ifelse(value %in% c("sentence  1|1") & varname %in% "do_timing_reorder" , "drop", value_normalized) ) %>%
                      filter(!is.na(value_normalized)) 
dim(reorder_normalized)

@@

```{r, eval=T}
# Fix typos:

print(paste("change 'premption' to 'preemption'", icb_long_spans_clean %>% dplyr::filter(stringr::str_detect(value, "premption")==TRUE) %>% nrow(), "times."))
icb_long_spans_clean <- icb_long_spans_clean %>% dplyr::mutate(value = stringr::str_replace(value, "premption", "preemption"))

print(paste("change 'disaprove' to 'disapprove'", icb_long_spans_clean %>% dplyr::filter(stringr::str_detect(value, "disaprove")==TRUE) %>% nrow(), "times."))
icb_long_spans_clean <- icb_long_spans_clean %>% dplyr::mutate(value = stringr::str_replace(value, "disaprove", "disapprove"))

print(paste("change 'uncoperative' to 'uncooperative'", icb_long_spans_clean %>% dplyr::filter(stringr::str_detect(value, "uncoperative")==TRUE) %>% nrow(), "times."))
icb_long_spans_clean <- icb_long_spans_clean %>% dplyr::mutate(value = stringr::str_replace(value, "uncoperative", "uncooperative"))

print(paste("change 'capitol' to 'capital'", icb_long_spans_clean %>% dplyr::filter(stringr::str_detect(value, "capitol")==TRUE) %>% nrow(), "times."))
icb_long_spans_clean <- icb_long_spans_clean %>% dplyr::mutate(value = stringr::str_replace(value, "capitol", "capital"))

print(paste("change 'survailance' to 'surveillance'", icb_long_spans_clean %>% dplyr::filter(stringr::str_detect(value, "survailance")==TRUE) %>% nrow(), "times."))
icb_long_spans_clean <- icb_long_spans_clean %>% dplyr::mutate(value = stringr::str_replace(value, "survailance", "surveillance"))

print(paste("change 'disaprove' to 'disapprove'", icb_long_spans_clean %>% dplyr::filter(stringr::str_detect(value, "disaprove")==TRUE) %>% nrow(), "times."))
icb_long_spans_clean <- icb_long_spans_clean %>% dplyr::mutate(value = stringr::str_replace(value, "disaprove", "disapprove"))

#Clean up some dates
icb_long_spans_clean <- icb_long_spans_clean %>% dplyr::mutate(value = stringr::str_replace(value, "17 july 2004", "17 July 2014"))

@@

@@

```{r message=FALSE}

target_file <- paste0(here::here(),"/replication_data/in/icb_manual_recoding_master_sheet.xlsx")
# in not working, run googledrive::drive_auth() and select new token

#Paper 2 will take dates and time seriously we ignore now
#What we're going to do is generate several new long variables and let it allow for agreements across coders
dictionary_dates <- readxl::read_excel(target_file, sheet="dates", col_types="text") %>%
                    dplyr::select(-coder, -notes)   %>% distinct() %>% 
                    tidyr::pivot_longer(cols=-c(date_orig)) %>% 
                    filter(!is.na(value)) 
dictionary_dates$decade <- ( dictionary_dates$value %>% stringi::stri_match_first(regex="([0-9]{3}[A-Z])") )[,2] 
dictionary_dates$year <- ( dictionary_dates$value %>% stringi::stri_match_first(regex="([0-9]{4})") )[,2] %>% as.numeric()
dictionary_dates$month <- ( dictionary_dates$value %>% stringi::stri_match_first(regex="[Y0-9]{4}-([0-9]{1,2})") )[,2] %>% as.numeric()
dictionary_dates$day <- ( dictionary_dates$value %>% stringi::stri_match_first(regex="[Y0-9]{4}-[M0-9]{1,2}-([0-9]{1,2})") )[,2] %>% as.numeric()

dictionary_dates <- dictionary_dates %>% 
                    mutate(month=ifelse(str_detect(value, "Jan|jan"),1, month)) %>%
                    mutate(month=ifelse(str_detect(value, "Feb|feb"), 2, month)) %>%
                    mutate(month=ifelse(str_detect(value, "March|mar"),3, month)) %>%
                    mutate(month=ifelse(str_detect(value, "April|apr"),4, month) )%>%
                    mutate(month=ifelse(str_detect(value, "May|may"),5, month) )%>%
                    mutate(month=ifelse(str_detect(value, "June|jun"),6, month)) %>%
                    mutate(month=ifelse(str_detect(value, "July|jul"),7, month) )%>%
                    mutate(month=ifelse(str_detect(value, "Aug|aug"),8, month) )%>%
                    mutate(month=ifelse(str_detect(value, "Sep|sep"),9, month) )%>%
                    mutate(month=ifelse(str_detect(value, "Oct|oct"),10, month) )%>%
                    mutate(month=ifelse(str_detect(value, "Nov|nov"),11, month) )%>%
                    mutate(month=ifelse(str_detect(value, "Dec|dec"),12, month) )%>%

                    #filter(value=="circa autumn 1947") %>%
                    mutate(decade = ifelse(!is.na(year), floor(as.numeric(year)/10)*10  ,  as.numeric(str_extract(decade, "[0-9]{3}"))*10  ) %>% as.numeric()  ) %>% 
                    mutate(year_part = NA ) %>%

                    mutate(year_part = ifelse( (!is.na(year) & is.na(month) & str_detect(value, "beginning|early|first half|second quarter|first quarter") ) | 
                                                 ( !is.na(month) & month<=4 ) , 1 , year_part ) ) %>%

                    mutate(year_part = ifelse( (!is.na(year) & is.na(month) & str_detect(value, "middle|summer|autumn") ) |  ( !is.na(month) & month>=5 & month<=8 ), 2 , year_part ) )  %>%

                    mutate(year_part = ifelse(  (!is.na(year) & is.na(month) & str_detect(value, "end") ) |  ( !is.na(month) & month>=9 )  , 3 , year_part )   ) %>%

                    mutate(month_part = NA ) %>%
                    mutate(month_part = ifelse( ( !is.na(year) & !is.na(month) & is.na(day) & str_detect(value, "beginning|early|first half")) |  ( !is.na(day) & day<=9 ) , 1 , month_part ) ) %>%
                    mutate(month_part = ifelse( (!is.na(year) & !is.na(month) & is.na(day) & str_detect(value, "middle")) |  (!is.na(day) &  day>=10 & day<=20 ), 2 , month_part ) )  %>%
                    mutate(month_part = ifelse( (!is.na(year) & !is.na(month) & is.na(day) & str_detect(value, "end")) |  (!is.na(day) &  day>=21 ), 3 , month_part ) ) %>%
                    dplyr::select(date_orig,  name, ,value , decade , year ,year_part , month , month_part, day ) %>%
                    arrange(decade,year, year_part, month, month_part, day) %>%
                    mutate(varname="date") %>%
                    mutate(varname_normalized="date" %>% paste0("_",name)) %>%
                    mutate(varname_normalized = ifelse(varname_normalized %in% c('date_earliest_date',
                                                                                 'date_earliest_date2',
                                                                                 'date_point_in_time',
                                                                                 'date_point_in_time2',
                                                                                 'date_point_in_time3',
                                                                                 'date_point_in_time4',
                                                                                 'date_point_in_time5',
                                                                                 'date_start_time',
                                                                                 'date_start_time2'
                                                                                 ) , 
                                                       "date_earliest" , varname_normalized   )) %>%
                    mutate(varname_normalized = ifelse(varname_normalized %in% c('date_earliest_end_date',
                                                                                 'date_end_time',
                                                                                 'date_end_time2',
                                                                                 'date_latest_date',
                                                                                 'date_latest_date2',
                                                                                 'date_latest_start_date'),  
                                                       "date_latest" , varname_normalized   ))

dictionary_dates %>% count(varname_normalized)

dates_normalized <- dictionary_dates %>% 
                    dplyr::select(date_orig, varname_normalized, decade , year, year_part, month, month_part,   day)  %>% 
                    pivot_longer(cols=-c(date_orig, varname_normalized)) %>%
                    mutate(varname_normalized=varname_normalized %>% paste0("_",name)) %>%
                    dplyr::select(value_normalized = value , value=date_orig, varname_normalized )

dictionary_actions  <- readxl::read_excel(target_file, sheet="actions") %>% dplyr::distinct() 
dictionary_actors    <- readxl::read_excel(target_file, sheet="actors") %>% dplyr::select(crisno,value, value_qcode=value_normalized_wikidata_id, value_normalized=value_normalized_label) %>% dplyr::distinct() %>%
                        full_join(varname_count %>% dplyr::select(varname) %>% filter(varname %>% str_detect("actor")),  by = character())

#Locations and dates are both really hard to do
dictionary_locations <- readxl::read_excel(target_file, sheet="locations", na = c("", "NA")) %>% 
                        dplyr::select(value=interact_location_other, value_normalized=norm_name, value_qcode= norm_qcode) %>% distinct() %>%
                        full_join(varname_count %>% dplyr::select(varname) %>% filter(varname %>% str_detect("location_other")),  by = character())
dictionary_condition_locations <- readxl::read_excel(target_file, sheet="condition_locations", na = c("", "NA")) %>% 
                                  dplyr::select(value=condition_interact_location_other, value_normalized=norm_name, value_qcode= norm_qcode) %>% distinct() %>%
                        full_join(varname_count %>% dplyr::select(varname) %>% filter(varname %>% str_detect("location_other")),  by = character())

dictionary_domains <- data.frame(
  stringsAsFactors = FALSE,
           varname = c("interact_domains",
                       "interact_domains","interact_domains","interact_domains",
                       "interact_domains","condition_interact_domains",
                       "condition_interact_domains","condition_interact_domains",
                       "condition_interact_domains","condition_interact_domains"),
             value = c("land","sea","air","wmd",
                       "space","land","sea","air","wmd","space"),
  value_normalized = c("land","sea","air","wmd",
                       "space","land","sea","air","wmd","space"),
       value_qcode = c("Q183366","Q15324","Q272730",
                       "Q49394","Q4169","Q183366","Q15324","Q272730",
                       "Q49394","Q4169")
)

dictionary_units <- data.frame(
  stringsAsFactors = FALSE,
                               varname = c("interact_units",
                                           "interact_units","interact_units",
                                           "interact_units","interact_units",
                                           "interact_units","interact_units",
                                           "interact_units","interact_units",
                                           "interact_units","interact_units",
                                           "interact_units","interact_units","interact_units",
                                           "interact_units",
                                           "condition_interact_units","condition_interact_units",
                                           "condition_interact_units",
                                           "condition_interact_units",
                                           "condition_interact_units",
                                           "condition_interact_units","condition_interact_units",
                                           "condition_interact_units",
                                           "condition_interact_units",
                                           "condition_interact_units","condition_interact_units",
                                           "condition_interact_units",
                                           "condition_interact_units",
                                           "condition_interact_units","condition_interact_units"),
                                 value = c("troops","armor",
                                           "artillery","fighters","bombers",
                                           "surface ships","missiles","nuclear",
                                           "aircraft carriers","surveillance",
                                           "submarines","survailance","chemical",
                                           "satellites","biological","troops",
                                           "armor","artillery","fighters",
                                           "bombers","surface ships","missiles",
                                           "nuclear","aircraft carriers",
                                           "surveillance","submarines","survailance",
                                           "chemical","satellites",
                                           "biological"),
                      value_normalized = c("troops","armor",
                                           "artillery","fighters","bombers",
                                           "surface ships","missiles","nuclear",
                                           "aircraft carriers","surveillance",
                                           "submarines","surveillance","chemical",
                                           "satellites","biological","troops",
                                           "armor","artillery","fighters",
                                           "bombers","surface ships","missiles",
                                           "nuclear","aircraft carriers",
                                           "surveillance","submarines",
                                           "surveillance","chemical","satellites",
                                           "biological"),
                           value_qcode = c("Q4991371","Q920182",
                                           "Q64418","Q127771","Q170877","Q11446",
                                           "Q974850","Q12802","Q17205",
                                           "Q3041792","Q2811","Q3041792","Q3639228",
                                           "Q26540","Q170907","Q4991371",
                                           "Q920182","Q64418","Q127771",
                                           "Q170877","Q11446","Q974850","Q12802",
                                           "Q17205","Q3041792","Q2811",
                                           "Q3041792","Q3639228","Q26540","Q170907")
                    )
